---
title: "Strings and factors"
output: html_document
date: "2023-10-17"
---

load the necessary libraries

```{r}
library(tidyverse)
library(p8105.datasets)
library(rvest)
library(httr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Strings and regex

detect a pattern in the strings - detect "jeff" in the strings
if we change jeff to Jeff, it will publish all False
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
```

detect a pattern in the strings - detect "e" in the strings
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "e")
```

Replace detected pattern to a new pattern
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
str_replace(string_vec, "jeff", "Jeff")
```

start with i think, but not end with i think, using `^` and `$`
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")
```

More complicated
capital or lower capital using `[]`
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )
str_detect(string_vec,"pumpkin")
str_detect(string_vec,"Pumpkin")
str_detect(string_vec,"[Pp]umpkin")
```

number first followed immediately by a letter - using `[]`
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

speical character 
detect 7 something in betwee and 11, using `.`
detect actual ., you need to put a \
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
str_detect(string_vec, "7\\.11")
```

speical character 
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

## Factors
string has ""
factor has levels 
```{r}
vec_sex = factor(c("male", "male", "female", "female"))

as.numeric(vec_sex)
```

what happends if i relevel ...
```{r}
vec_sex = fct_relevel(vec_sex, "male")

as.numeric(vec_sex)
```

## NSDUH -- Strings

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) |> 
  html_table() |> 
  first() |>
  slice(-1)
```

```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) |>
  separate(age_year, into = c("age", "year"), sep = "\\(")|>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)
  )|>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```


## NSDUH -- factors

```{r}
data_marj |>
  filter(age == "12-17") |>
  mutate(State =fct_relevel(State, "Texas", "Oklahoma"))|>
  ggplot(aes(State, y = percent, color = year)) +
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
data_marj |>
  filter(age == "12-17") |>
  mutate(State =fct_reorder(State, percent))|>
  ggplot(aes(State, y = percent, color = year)) +
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```










